Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''

### 2023-02-06 additional_files/additional_file_1.tex
Add main text contents to the repository; manuscript.tex file includes both manuscript and supplemental data into one pdf

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc} % declare encoding as utf8
\usepackage{graphicx} % to enable \includegraphics
\usepackage{flafter} % to make sure floats appear after their position in text
\usepackage{pdflscape} % allow some pages to be in landscape orientation
\usepackage{layout} % for debugging purposes, looking at page layout
\usepackage[margin=1in]{geometry} % leaves more space for text body
\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
\usepackage{csvsimple} % read and display csv files as tables
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables
\usepackage{textcomp} % makes the \textdegree command available
\usepackage{tocloft} % formatting the lists of tables and figures
\usepackage{xcolor} % enable text coloring
\usepackage{calc} % makes \widthof available to set the width of the last column
\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
\usepackage{longtable} % allow tables to span more than one page
\usepackage{threeparttablex} % threeparttable for longtable
\usepackage{capt-of} % put caption on next page
\usepackage{listings}
\usepackage{datatool} % allows fetching key-value pairs from a text file
\usepackage{natbib}

% Setting the separator to use with datatool
\DTLsetseparator{ = }

% Formatting the table of contents/subsection
\renewcommand{\contentsname}{Supplemental methods}

% Formatting the list of tables
\renewcommand{\listtablename}{Supplemental tables} % Write "Supplemental tables" instead of "List of tables"
\renewcommand\cfttabpresnum{Table } % Write "Table" before table number in list of tables
\setlength{\cfttabnumwidth}{6em} % Spacing before captions
\setlength{\cftbeforefigskip}{1ex} % Vertical spacing between entries

% Formatting the list of figures
\renewcommand{\listfigurename}{Supplemental figures} % Write "Supplemental figures" instead of "List of figures"
\renewcommand\cftfigpresnum{Figure } % Write "Figure" before fig number in list of figures
\setlength{\cftfignumwidth}{6em} % Spacing before captions
\setlength{\cftbeforetabskip}{1ex} % Vertical spacing between entries

\graphicspath{ {../figures/} } % tell Latex to look for figures in figures/

% Format figure and table numbers to follow S1, S2, ...
\renewcommand\thefigure{S\arabic{figure}} 
\renewcommand\thetable{S\arabic{table}}

% Defining new environments as I will always want the figures centered
\newenvironment{cfigure}
	{\begin{figure} \centering}
	{\end{figure}}

\newenvironment{lsfigure}
	{\begin{landscape} \begin{figure} \centering}
	{\end{figure} \end{landscape}}

\setcounter{totalnumber}{1} % Allow a maximum of one float (table/figure) per page

\setlength{\parskip}{2ex}

% Defining a command to generate a table of phenotypic data for a single gwas
\newcommand{\gwastableone}[2]{
\begin{table}
	\caption[Phenotypic data used for the GWAS analysis of #2]%
	{Phenotype frequency and numerical coding of phenotypic data used for the GWAS analysis of #2.
	``NA'' indicates a missing value in the source data.
	Dashes indicate that this phenotype was not used in the GWAS analysis.}
	\label{#1_gwas_table}

	\centering
	\begin{threeparttable}\small

		\csvreader[head to column names,
		/csv/separator=semicolon,
		tabular = llc,
		table head = \toprule Phenotype & Frequency & GWAS \\ \midrule,
		table foot = \bottomrule]%
		{../tables/#1_gwas_table.csv}%
		{}%
		{\Value & \Frequency & \GWASa}

	\end{threeparttable}

\end{table}
\clearpage
}

% Defining a command to generate a table of phenotypic data for two gwas
\newcommand{\gwastabletwo}[2]{
\begin{table}
	\caption[Phenotypic data used for the GWAS analyses of #2]%
	{Phenotype frequency and numerical coding of phenotypic data used in two GWAS analyses of #2.
	``NA'' indicates a missing value in the source data.
	Dashes indicate that this phenotype was not used in the GWAS analysis.}
	\label{#1_gwas_table}

	\centering
	\begin{threeparttable}\small

		\csvreader[head to column names,
		/csv/separator=semicolon,
		tabular = llcc,
		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2\\ \midrule,
		table foot = \bottomrule]%
		{../tables/#1_gwas_table.csv}%
		{}%
		{\Value & \Frequency & \GWASa & \GWASb}

	\end{threeparttable}

\end{table}
\clearpage
}

% Defining a command to generate a table of phenotypic data for three gwas
\newcommand{\gwastablethree}[2]{
\begin{table}
	\caption[Phenotypic data used for the GWAS analyses of #2]%
	{Phenotype frequency and numerical coding of phenotypic data used in three GWAS analyses of #2.
	``NA'' indicates a missing value in the source data.
	Dashes indicate that this phenotype was not used in the GWAS analysis.}
	\label{#1_gwas_table}

	\centering
	\begin{threeparttable}\small

		\csvreader[head to column names,
		/csv/separator=semicolon,
		tabular = llccc,
		table head = \toprule Phenotype & Frequency & GWAS \#1 & GWAS \# 2 & GWAS \# 3\\ \midrule,
		table foot = \bottomrule]%
		{../tables/#1_gwas_table.csv}%
		{}%
		{\Value & \Frequency & \GWASa & \GWASb & \GWASc}

	\end{threeparttable}

\end{table}
\clearpage
}

% Defining a command for manhattan plot figures
\newcommand{\manhattanplot}[3]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_manhattan}
	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	The x-axis shows the position along the reference assembly version 4 of Williams82.
	Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	Vertical dotted lines indicate the position of signals associated with the trait.
	Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
	The ``Gm'' prefix has been left out of chromosome names for simplicity.
	}
	\label{#1_manhattan}
\end{cfigure}

\clearpage}

% Defining a new command for signal plot figures
\newcommand{\signalplot}[4]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_signal}
	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	Panel (a) shows gene models over the genomic interval.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	#4
	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	}
	\label{#1_signal}
\end{cfigure}

\clearpage}

% Defining a new command for gene plot figures
\newcommand{\geneplot}[5]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_gene}
	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	#5
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	Panel (a) shows gene models over the plotting interval.
	Exons are represented by rectangles whereas introns are represented by horizontal lines.
	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
	}
	\label{#1_gene}
\end{cfigure}

\clearpage}

% Defining a new command for k-mer plot figures
\newcommand{\kmerplot}[5]{
	\begin{cfigure}
		\includegraphics[width = 5.8in]{#1_kmers}
		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
		The red rectangle identifies the region that is zoomed-in in panel (b).
		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
		\label{#1_kmers}
	\end{cfigure}

\clearpage}

% Defining a new command for LD plot figures
\newcommand{\ldplot}[3]{
	\begin{cfigure}
		\includegraphics[width = 5.8in]{#1_ld}
		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
		as identified by ``Gm'' chromosome labels.
		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
		\emph{k}-mers are represented in the same order along the x- and y-axis.
		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
		\label{#1_ld}
	\end{cfigure}

\clearpage}

% Information for the title page
\title{Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''}
\author{Marc-André Lemay \and Jonas A. Sibbesen \and Maxime de Ronne \and Richard Bélanger \and François Belzile}
\date{}

% END OF PREAMBLE

\begin{document}

@@

% Detailed methods for some aspects of the paper
\maketitle \thispagestyle{empty}

\listoftables
\listoffigures

\thispagestyle{empty}

\section{Detailed methods}
\label{annexe-sv-gwas-detailed-methods}

@@

\end{document}

>>>>

\chapter{Supplemental data}
\label{supdata}

@@

% Detailed methods for some aspects of the paper
\section{Detailed methods}
\label{annexe-sv-gwas-detailed-methods}

@@
